---
title: "574_HW3"
author: "Matthew Stoebe"
date: "2025-04-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1


## a&b)
```{r}


oil <- read.table("Data/oil.dat", header = FALSE)
colnames(oil) <- c("X1", "X2", "X3", "X4", "X5", "zone")


groups <- unique(oil$zone)
g <- length(groups)
n_total <- nrow(oil)

xbar <- colMeans(oil[, c("X1","X2","X3","X4","X5")])

p <- 5
B <- matrix(0, nrow = p, ncol = p)
W <- matrix(0, nrow = p, ncol = p)


for (grp in groups) {

  oil.grp <- subset(oil, zone == grp)
  n_grp <- nrow(oil.grp)
  

  xbar_grp <- colMeans(oil.grp[, c("X1", "X2", "X3", "X4", "X5")])
  
  diff_grp <- as.numeric(xbar_grp - xbar)
  B <- B + n_grp * (diff_grp %*% t(diff_grp))
  
  for (i in 1:n_grp) {
    diff_i <- as.numeric(oil.grp[i, c("X1", "X2", "X3", "X4", "X5")] - xbar_grp)

    W <- W + outer(diff_i, diff_i)
  }
}


A <- solve(W) %*% B


eigenA <- eigen(A)$values
print("Eigenvalues of A:")
print(eigenA)



Lambda <- Re(prod(1 / (1 + eigenA)))
print("Wilks' Lambda:")
print(Lambda)


T_stat <- - (n_total - 1 - p + g/2) * log(Lambda)
df <- p * (g - 1)
p_value <- 1 - pchisq(T_stat, df = df)
cat("Test Statistic T =", T_stat)


```
## c)
```{r}
fit <- manova(cbind(X1, X2, X3, X4, X5) ~ zone, data=oil)
summary_fit <- summary(fit, test="Wilks")
print(summary_fit)

```
## d)
```{r}
summary_aov <- summary.aov(fit)
print(summary_aov)

```

## e)
```{r}
fit12 <- manova(cbind(X1, X2, X3, X4, X5) ~ I(zone %in% c("Wilhelm","Sub-Muhinina")), data=oil)
summary(fit12, test="Wilks")

```
# Question 2



```{r}
stock <- read.table("Data/stock.dat", header = FALSE)

n <- nrow(stock)
p <- ncol(stock)


pca_stock <- prcomp(stock, scale. = TRUE)
summary(pca_stock)
pca_stock$rotation
head(pca_stock$x)
```

## a)
```{r}
variance_Y2 <- pca_stock$sdev[2]^2
cat("Estimated variance of Y2 =", variance_Y2, "\n\n")
```

## b)
```{r}
xi <- c(0.58, -0.41, -0.32, -1.82, 0.04)

loading1 <- pca_stock$rotation[, 1]
Y1_value <- sum(xi * loading1)
cat("The first principal component (Y1) for xi is", Y1_value, "\n\n")
```

## c)

```{r}
summary(pca_stock)

# Create a scree plot:
plot(pca_stock, type = "l", main = "Scree Plot of Stock Returns PCA")
```

## d)
```{r}
loadings <- pca_stock$rotation
cat("Loadings for PC1:\n")
print(loadings[, 1])
cat("\nLoadings for PC2:\n")
print(loadings[, 2])
cat("\n")
```
## e)

```{r}
pc_scores <- pca_stock$x

plot(pc_scores[, 1], pc_scores[, 2],
     xlab = "PC1", ylab = "PC2",
     main = "PCA of Stock Returns (Weeks)",
     pch = 19, col = "gray")

text(pc_scores[, 1], pc_scores[, 2], labels = 1:n, cex = 0.7, pos = 4)

points(pc_scores[63, 1], pc_scores[63, 2], col = "red", pch = 19, cex = 1.5)
text(pc_scores[63, 1], pc_scores[63, 2], labels = "63", col = "red", cex = 0.9, pos = 3)
```

It appears that week 63 is an extreme outlier for PC1 and about average for PC2.

# Question 3
```{r}
S <- matrix(c(
  6.59, 6.87, 7.01, 6.56, 6.67, 6.83, 5.97, 7.12,
  6.87, 9.79, 10.79, 10.72, 10.42, 11.01, 11.36, 13.44,
  7.01, 10.79, 13.97, 14.63, 14.13, 14.57, 15.77, 18.58,
  6.56, 10.72, 14.63, 16.72, 16.41, 16.99, 18.82, 21.76,
  6.67, 10.42, 14.13, 16.41, 17.37, 18.17, 19.62, 22.33,
  6.83, 11.01, 14.57, 16.99, 18.17, 21.52, 25.79, 29.68,
  5.97, 11.36, 15.77, 18.82, 19.62, 25.79, 39.47, 48.76,
  7.12, 13.44, 18.58, 21.76, 22.33, 29.68, 48.76, 75.14),
  nrow = 8, ncol = 8, byrow = TRUE)
print("Covariance matrix S:")
print(S)

mean_vec <- c(5.96, 9.07, 12.18, 14.89, 18.26, 20.47, 20.13, 14.59)

lambda1 <- 159.28
lambda2 <- 28.54
lambda3 <- 6.59

e1 <- c(-0.11, -0.18, -0.24, -0.28, -0.28, -0.34, -0.47, -0.63)
e2 <- c(-0.27, -0.32, -0.37, -0.36, -0.36, -0.22,  0.16,  0.59)
e3 <- c( 0.46,  0.39,  0.21, -0.05, -0.19, -0.36, -0.50,  0.41)
```


## a)
We can use the covariance matrix here becuase all fields are measured in the same units (snow depth) so the variance values are meaningful relative to one another. 

## b)
159.28

## c)
0

## d)
10.79

## e)
```{r}
cov_Y1_X1 <- lambda1 * e1[1]
cov_Y1_X1
```

## f)
```{r}
total_variability <- sum(diag(S))
explained_variability_PC1_PC2 <- lambda1 + lambda2
prop_explained <- explained_variability_PC1_PC2 / total_variability
round(prop_explained * 100, 2)
```

## g)
PC1 has consistent direction and magnitude. This indicates that PC1 reflects overall level of the snow pack where end of season periods contribute heavily. Because it is all negative, negative 

## h)
PC2 has negative loading for early periods and positive for later periods which contrasts early snow accumulation with late season accumulation

## i)
1991s overall snowpack was average, but the high PC2 indicates that the majority of the snow was late season

## j)
It looks like 2011 was the best year for rafting as there is a negative pc1 (negative pc1 means large snow pack i think) and a high PC2 which means a lot of the snow came later in the year

## k)
```{r}
y8 <- c(-7.50, -2.25, -4.25)
x8_centered <- y8[1]*e1 + y8[2]*e2 + y8[3]*e3
x8_reconstructed <- mean_vec + x8_centered
round(x8_reconstructed, 2)
```

## l)
```{r}
x1 <- c(1.53, 2.47, 2.76, 4.76, 7.32, 11.20, 14.55, 3.11)
x1_centered <- x1 - mean_vec
y1_2 <- sum(e2 * x1_centered)
round(y1_2, 2)

```

