---
title: "574_HW5"
output: pdf_document
---

# Question 1


## a
```{r}
library(MASS)
library(ggplot2)

root <- read.table("Data/rootstock.txt", header = TRUE, stringsAsFactors = FALSE)

root$RootStock <- factor(root$RootStock)
str(root)

lda_fit <- lda(RootStock ~ Y1 + Y2 + Y3 + Y4, data = root)

lda_fit$scaling

```
## b
```{r}
prop_var <- lda_fit$svd^2 / sum(lda_fit$svd^2)
round(prop_var, 3)  
cumsum(prop_var)
```
Because DF1 + DF2 explain 91 % of the between-group variance, I retain two discriminant functions.

## c
```{r}

scores <- as.data.frame(predict(lda_fit)$x) 
scores$RootStock <- root$RootStock   
ggplot(scores, aes(LD1, LD2, colour = RootStock)) +
  geom_point(size = 2) +
  labs(title = "Root-stock study: LD1 vs LD2") +
  theme_bw()

```

Stocks 1 and 6 cluster on the left of LD1 and around the midpoint of LD2. Stocks 2 and 3 cluster on the right of LD1 and spread across most of LD2. These two seem well separated. Stocks 4 and 5 are more mixed in with the rest and more difficult to separate.

## d
```{r}

new_tree <- data.frame(Y1 = 1.09, Y2 = 2.69, Y3 = 4.38, Y4 = 1.29)

pred <- predict(lda_fit, new_tree)
pred$class
pred$posterior
```
Rootstock 3 has the largest posterior probability, but the model is not super confident between stocks 3 and 5 both having a probability around .3


# Question 2
Basic Data
```{r}
a <- c(0.32, 0.05, 0.19, -2.13, -0.14)

bird1 <- c(156, 245, 31.60, 18.50, 20.50) 

mean_surv <- c(157.3, 241.0, 31.4, 18.5, 20.8)
mean_died <- c(158.4, 241.5, 31.4, 18.4, 20.8)

pi <- c(surv = 0.43, died = 0.57)
```

## a

```{r}
Z  <- sum(a * bird1)    
cat("Bird 1 Z score is", Z)
```

## b

```{r}
Z_surv <- sum(a * mean_surv)
Z_died <- sum(a * mean_died)

cat("Survived mean is", Z_surv, " \nDied mean is", Z_died)

```

## c
```{r}
midpt   <- 0.5 * (Z_surv + Z_died)          # 26.33
class_eq <- ifelse(Z < midpt, "survived", "died")

cat("The Midpoint is", midpt, "\nGiven this data, Bird 1 ", class_eq)
```

## d

The Probaibliteis provided classify the bird as dead, but differs from the midpoint rule where the bird survives. This is because predict uses a larger prior for "died" which shifts the decision boundary.

## e
We assume that the sample rates accurately represent the true population rates for sparrows that strike windows

## f
No. LDA uses the pooled covariance matrix and effectively rescales each variable naturally.
