---
title: "573_HW_5"
author: "Matthew Stoebe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

install.packages("astsa")

```

#Question 1

```{r}
library(MARSS)
library(astsa) # Contains the global temperature data

# Load the data
library(MARSS)
library(astsa) # this package contains the data
Y <- cbind(xglobtemp,xglobtempl)
ts.plot(Y, col=c(6,4), ylab='Temperature Deviations')
legend("topleft", legend = c("Land & Ocean", "Land"), col = c(6, 4), lty = 1)

time <- time(Y)
```
##a
```{r}
# Define the model list
model.list <- list(
  Z = matrix(1, nrow = 2, ncol = 1),                      
  B = matrix(1),                                           
  U = matrix("delta"),                                     
  Q = matrix("q"),                                         
  R = matrix(list("r11", "r12", "r12", "r22"), nrow = 2),  
  x0 = matrix(-0.35),                                      
  V0 = matrix(1)                                           
)

# Fit the model
fit <- MARSS(Y, model = model.list, method = "BFGS")

logLik(fit)
fit$AICc
residuals <- residuals(fit)


```
##b
What are the unknown parameters? What are your estimates of the unknown parameters? Write down
the fitted model. (The estimates of the variance parameters do not match the book, because the book
returns estimates of the cholesky decomposition chol(R) instead of the variance-covariance matrix R.)
```{r}
coef_estimates <- coef(fit, type = "vector")
print(coef_estimates)
```

##c
What are the standard errors of your estimates in (b). Hint: use MARSSparamCIs function.
```{r}

param_CIs <- MARSSparamCIs(fit)
print(param_CIs$par.se)

```
##d 
```{r}
states <- fit$states

time_sequence <- seq_along(Y[, 1])  
plot(time_sequence, states[1, ], type = "l", col = "red", ylab = "Temperature Deviations", 
     xlab = "Time", main = "Smoothed Global Temperature Anomalies", lwd = 5)
lines(time_sequence, Y[, 1], col = 6, lty = 2)
lines(time_sequence, Y[, 2], col = 4, lty = 2)
legend("topleft", legend = c("Smoothed State", "Land & Ocean", "Land"), 
       col = c("red", 6, 4), lty = c(1, 2, 2))
```
##e

Unconstrained R Assumptions
- Errors vt1 and vt2 can have different variances.
- There may be nonzero covariance between the errors

Diagonal and Unequal R Assumptions
- Errors Vt1 and vt2 can have different variances
- There is no covariance between the errors => Observation errors are uncorrelated

Diagonal and Equal R:
- Errors are uncorrelated and have the same variance

##f
Fit the two models in (e). Compare the AIC values of the three different models for R, which one is
better?
```{r}

fit_unconstrained <- fit  # From part (b)

model_diag_unequal <- model.list
model_diag_unequal$R <- matrix(list("r11", 0, 0, "r22"), 2, 2)
fit_diag_unequal <- MARSS(Y, model = model_diag_unequal, method = "BFGS")

model_diag_equal <- model.list
model_diag_equal$R <- matrix(list("r11", 0, 0, "r11"), 2, 2)
fit_diag_equal <- MARSS(Y, model = model_diag_equal, method = "BFGS")

AIC_values <- data.frame(
  Model = c("Unconstrained R", "Diagonal and Unequal R", "Diagonal and Equal R"),
  AICc = c(fit_unconstrained$AICc, fit_diag_unequal$AICc, fit_diag_equal$AICc)
)

print(AIC_values)

```

#Question 2
Modify the model above to make it slightly more flexible. We keep the state equation the same as problem 1,
since the random walk model with drift makes sense for this data set as we have seen in the previous lecture.
However, we will modify the observation equation to allow for different coefficients (a1, a2) for the two series,
and we will estimate the initial state parameters x0 ∼ N(μ0, σ2) instead of fixing it.

##a

```{r}


Y <- cbind(xglobtemp,xglobtempl)

# Plot the data to visualize
ts.plot(Y, col = c(6, 4), ylab='Temperature Deviations', main = 'Global Temperature Deviations')
legend("topleft", legend = c("Land & Ocean", "Land"), col = c(6, 4), lty = 1)

dim(Y)

model_modified <- list(
  Z = matrix(c("a1", "a2"), nrow = 2, ncol = 1),             # Observation matrix with estimated coefficients
  A = matrix(0, nrow = 2, ncol = 1),                         # Observation intercept matrix set to zero
  B = matrix(1),                                             # State transition matrix (random walk)
  U = matrix("delta"),                                       # State intercept (to estimate)
  Q = matrix("q"),                                           # State noise variance (to estimate)
  R = matrix(list("r11", "r12", "r12", "r22"), nrow = 2, ncol = 2),    # Observation noise covariance matrix
  x0 = matrix("x0", nrow = 1, ncol = 1),                     # Initial state mean (to estimate)
  V0 = matrix("V0", nrow = 1, ncol = 1)                      # Initial state covariance (to estimate)
)

fit_modified <- MARSS(Y, model = model_modified, method = "BFGS", 
                      control = list(maxit = 500)) # Increased max iterations if needed
```


```{r}
MARSSinfo()
```

