---
title: "573_FinalExam"
author: "Matthew Stoebe"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## a 
Which plot (Left or Right) of the LJung-box statistics suggests autocorrelation exists in the time series.Justify your answer. (2 pts)

The left  plot indicates that there is autocorrelation as the different lag values have signifficance below the p value line. 

## b
What are the null and alternative hypotheses corresponding to the p-value the arrow points to? (3 pts)

- Null Hypothesis: there is no autocorrelation up to a a specific lag
- Alternative hypothesis: there is some autocorrelation at one or more legs

# Question 2
Im assuming here that "quantities" refer to either predicting, filtering, smoothing, and forecasting.It seems that the quantities of interest are defined in the question itself?

## a
current position and velocity would be a filtering problem

## c (why is this out of order?)
Predicting during the huricane wold be a forecasting problem

## b 
this would be a smoothing operation


# Question 3

Fit HHMM model with two hidden states 

```{r}
library(depmixS4)
library(astsa)

data(qinfl)

mod <- depmix(response = qinfl ~ 1, 
              data = data.frame(qinfl=qinfl), 
              nstates = 2, 
              family = gaussian())

fitmod <- fit(mod)

summary(fitmod)

```

# Fit linear Gaussian state-space model 
```{r}
library(MARSS)
inflation <- qinfl

Y <- matrix(inflation, nrow = 1)

model.list <- list(
  Z = matrix(1),
  A = matrix(0),
  B = matrix("b1"),
  U = matrix("u1"),
  Q = matrix("q1"),
  R = matrix("r1"),
  x0 = matrix("x0"),
  V0 = matrix("v0")
)


fit_initial <- MARSS(Y, model = model.list, method = "kem")
fit_final <- MARSS(Y, model = model.list, inits = fit_initial, method = "BFGS")


summary(fit_final)

states = fit_final$states
time_index <- seq_along(qinfl)  

states <- as.numeric(fit_final$states)
qinfl_data <- as.numeric(qinfl)

plot.ts(qinfl_data, type = "l", col = "black", lwd = 2,
        ylab = "Rate",
        main = "Observed Inflation vs. Smoothed State")
lines(states, col = "blue", lwd = 2)

legend("topleft", legend = c("Observed Inflation", "Smoothed State"),
       col = c("black", "blue"), lwd = 2)


```
This is similar to stochastic regression as it allows for the relationships to vary with time instead of being set. 

## c
HMM assumes that the process swaps between discrete spaces with specific feature sets while gaussian state assumes that the evolution is continual. 

## d
Stochastic Regression
```{r}
data(qintr)

len <- length(qinfl)
Z <- array(NA, c(1, 2, len))
Z[1, 1, ] <- 1
Z[1, 2, ] <- qintr

model_list <- list(
  B = diag(2),
  U = "zero",
  A = matrix(0,1,1),
  Q = "diagonal and unequal",
  Z = Z,
  R = matrix("r",1,1),
  x0 = matrix(c("x01","x02"), 2, 1),
  V0 = diag(0.01, 2),
  tinitx = 0
)

inits <- list(
  x0 = matrix(c(0, 0), 2, 1)
)

fit_init <- MARSS(qinfl, model=model_list, method="kem", inits=inits)
fit_final <- MARSS(qinfl, model=model_list, method="BFGS", inits=fit_init)

summary(fit_final)

```

## e
in model d, inflation is described by another observed variable with a time varying slope. this is not available in model b. 

## f
This really depends on how you view inflation and its relation to time. I would personally choose the final model as it has a better view of the general macroeconomic enviroment with the incorporation of interest rate.