---
title: "STAA 552: HW 1"
author: "Matthew Stoebe"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
    includes:
      in_header: preamble.tex
header-includes:
- \usepackage[default]{sourcesanspro}
- \usepackage[T1]{fontenc}
mainfont: SourceSansPro
---

```{r setup, include=FALSE}
#Retain this code chunk!!!
library(knitr)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

See Canvas Calendar for due date.  
52 points total, 4 pts per problem unless otherwise noted.   
Content for Q1-Q7 is from section 01.  
Content for Q8-Q14 is from section 02.  
Add or delete code chunks as needed.  
For full credit, your numeric answers should be clearly labeled, outside of the R output.

# Q1 - Q3

For this group of questions, identify each variable as **nominal, ordinal or numeric**. 


:::: {.taskbox data-latex=""}
## Q1 (2 pts)
Anxiety rating (none, mild, moderate, severe)
::::

*****
Response

Ordinal

*****

:::: {.taskbox data-latex=""}
## Q2 (2 pts)
Favorite grocery store (Safeway, King Soopers, Whole Foods, other)
::::

*****
Response
Nominal
*****

:::: {.taskbox data-latex=""}
## Q3 (2 pts)
Annual income ($)
::::


*****
Response
Numeric
*****

# Poisson Distribution (Q4 - Q6)

Note: These questions are "self-checking" because the theoretical and simulated distributions should yield similar results. 

:::: {.taskbox data-latex=""}
## Q4

Simulate 5000 independent replicates of the random variable Y $\sim$ Poisson($\mu$) with $\mu$ = 7.5. Use set.seed(5821) for reproducibility. Calculate the sample mean and sample variance.  
::::


*****
Response

```{r}
#Q4
set.seed(5821)

n = 5000

Y = rpois(n, lambda = 7.5)

sample_mean <- mean(Y)
sample_variance = var(Y)

cat("sample Mean:", sample_mean, "\n")
cat("Sample Variance:", sample_variance, "\n")


```

*****

:::: {.taskbox data-latex=""}
## Q5

Plot the empirical probability mass function (ex: density histogram) of your simulated values.  Overlay the true probability mass function for Poisson(7.5) in red.  

Note: Some example R code has been provided to get you started.  But other approaches are allowed.
::::


```{r}
#Q5
x = seq(0, max(Y))
hist(Y, freq = FALSE, breaks = seq(-0.5, max(Y)+0.5, 1))
#This choice of breaks is suggested to center each bar at an integer value since the Poisson distribution takes only integer values!
points(x, dpois(x, 7.5), col = "red", pch = 16)
```

:::: {.taskbox data-latex=""}
## Q6

Plot the empirical probability mass function (ex: density histogram) of your simulated values.   Overlay the normal approximation with mean = 7.5 and variance = 7.5.  

Notes: Since $\mu$ < 10 we don't expect a great fit. Some example R code has been provided to get you started.  But other approaches are allowed.
::::



```{r}
#Q6
x = seq(0, max(Y))
hist(Y, freq = FALSE, breaks = seq(-0.5, max(Y)+0.5, 1))
#This choice of breaks is suggested to center each bar at an integer value since the Poisson distribution takes only integer values!
points(x, dpois(x, 7.5), col = "red", pch = 16)
#This choice of breaks is suggested to center each bar at an integer value since the Poisson distribution takes only integer values!
curve(dnorm(x, mean = 7.5, sd = sqrt(7.5)), col="blue", lwd=2, add = TRUE)
```


# Diabetes (Q7 - Q10)

Suppose it is known that 12\% of US adults have diabetes.  Hence, $\pi$ = 0.12. Consider a random sample of n = 160 US adults.  Let Y represent the number of people with diabetes (in a given sample).  Let  $\hat\pi = (Y/n)$ represent the sample proportion of people with diabetes.  

Note: Q8-Q10 are "self-checking" because the theoretical and simulated distributions should yield similar results. 

:::: {.taskbox data-latex=""}
## Q7 (6 pts)

Specify the (exact) distribution for Y.  Give E(Y) and Var(Y).
::::


*****
Response

This is a Binomial Distribution with p = .12 (probability of selecting a person with diabetes) and n = 160

```{r}

#Q7
n = 160
p = .12

E_Y <- n*p
Var_y <- n*p*(1-p)

cat("Expected Value:", E_Y, "\n")
cat("Expected Variance:", Var_y, "\n")


```

*****

:::: {.taskbox data-latex=""}
## Q8 (6 pts)

Specify the (approximate) distribution for $\hat\pi$.  Give E($\hat\pi$) and Var($\hat\pi$).
::::


*****
Response

The Binominal Distribution can be approximated by the Normal distribution when N is large and p is small.
```{r}
#Q8
n = 160
p = .12

E_pihat <- p
Var_pihat <- p*(1-p)/n

cat("Expected Value:", E_pihat, "\n")
cat("Expected Variance:", Var_pihat, "\n")


```

*****

:::: {.taskbox data-latex=""}
## Q9

Simulate 5000 independent replicates from the distribution specified in Q7. Use set.seed(4966) for reproducibility. Then calculate the corresponding $\hat\pi$ values.  Calculate the sample mean and sample variance of the $\hat\pi$ values.  
::::


*****
Response

```{r}
#Q9
set.seed(4966)

i = 5000


Y = rbinom(i, n,p)

sample_mean <- mean(Y)
sample_variance = var(Y)

pihat = Y/n

pihat_mean <- sample_mean / n
pihat_variance <- sample_variance/n**2

cat("sample Mean:", pihat_mean, "\n")
cat("Sample Variance:", pihat_variance, "\n")

```

*****

:::: {.taskbox data-latex=""}
## Q10

Plot the empirical distribution (ex: density histogram) of your observed $\hat\pi$ values from the previous question.  Overlay the approximate distribution from Q8.  
::::


```{r}
#Q10

hist(pihat, freq = FALSE, breaks = 50, 
     main = "Empirical Distribution of pihat and Approximate Normal Distribution",
     xlab = "pihat", col = "lightblue", border = "black")

x_vals <- seq(min(pihat), max(pihat), length = 100)
y_vals <- dnorm(x_vals, mean = E_pihat, sd = sqrt(Var_pihat))
lines(x_vals, y_vals, col = "red", lwd = 2)


```


# Male Births (Q11 - Q14)

Let $\pi$ be the true proportion of babies that are male in some large population.  Suppose we have a random sample of n = 1574 birth records this population.  Of these, y = 803 are male.   We will use this data to test $H_0: \pi = 0.5$ vs $H_A: \pi \ne 0.5$.  

Note: Q12-Q14 are "self-checking" because all three methods will give very similar results. 

:::: {.taskbox data-latex=""}
## Q11 (2 pts)

Calculate $\hat\pi$.
::::


*****
Response

```{r}
#Q11
n <- 1574
y <- 803
pi <- .5

pihat <- y/n

cat("observed rate pihat:", pihat, "\n")

```

*****

:::: {.taskbox data-latex=""}
## Q12 

Run the test of interest using a **Wald** test.  Report the $\chi^2$ test statistic and corresponding p-value.   Do this "by hand" (using R as a calculator) and echo your R code.
::::

*****
Response

```{r, echo = TRUE}
#Q12

pi0 <- .5

z <- (pihat - pi0) / sqrt(pihat * (1 - pihat)/n)

x <- z**2

p <- pchisq(x, df=1, lower.tail = FALSE)

cat("wald test chi squared statistic is:", x, "\n")
cat("with a p value of", p, "\n")

```

*****

:::: {.taskbox data-latex=""}
## Q13

Run the test of interest using a **Score** test.  Report the $\chi^2$ test statistic and corresponding p-value.   Do this "by hand" (using R as a calculator) and echo your R code.
::::

*****
Response

```{r, echo = TRUE}
#Q13
pi0 <- .5

z <- (pihat - pi0) / sqrt(pi0 * (1 - pi0)/n)

x <- z**2

p <- pchisq(x, df=1, lower.tail = FALSE)

cat("score test chi squared statistic is:", x, "\n")
cat("with a p value of", p, "\n")
```

*****

:::: {.taskbox data-latex=""}
## Q14

Run the test of interest using a **likelihood ratio** test. (See notes or CDA 1.4.1 for details and formula for test statistic.) Report the $\chi^2$ test statistic and corresponding p-value.   Do this "by hand" (using R as a calculator) and echo your R code.
::::

*****
Response

```{r, echo = TRUE}
#Q14

L0 <- y * log(pi0) + (n - y) * log(1 - pi0)

L1 <- y * log(pihat) + (n - y) * log(1 - pihat)


likelyhood_ratio = -2*(L0 - L1)


p <- pchisq(likelyhood_ratio, df = 1, lower.tail = FALSE)


cat("likelyhood_ratio test chi squared statistic is:", likelyhood_ratio, "\n")
cat("with a p value of", p, "\n")

```

*****

# Appendix
```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}
```