---
title: "STAA 552: HW 2"
author: "YOUR NAME HERE"
output:
  pdf_document:
    toc: no
    toc_depth: '2'
    includes:
      in_header: preamble.tex
header-includes:
- \usepackage[default]{sourcesanspro}
- \usepackage[T1]{fontenc}
mainfont: SourceSansPro
---

```{r setup, include=FALSE}
#Retain this code chunk!!!
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

See Canvas Calendar for due date.  
54 points total.  
Content for Q1-Q8 is from section 02.  
Content for Q9-Q14 is from section 03.  
Add or delete code chunks as needed.  
For full credit, your numeric answers should be clearly labeled, outside of the R output.

# Diabetes 2 (Q1 - Q6)

A researcher is interested in estimating the proportion of US adults who have been diagnosed with diabetes ($\pi$).  Consider a study where n = 65 people are contacted and asked whether they have been diagnosed with diabetes.   Suppose this is based on a random sample of US adults.  From this sample, 7 people have been diagnosed with diabetes.

:::: {.taskbox data-latex=""}
## Q1 (2 pts)

Calculate $\hat\pi$.
::::

*****
Response

```{r}
#Q1

```

*****

:::: {.taskbox data-latex=""}
## Q2 (4 pts)

Discuss whether the large sample normal approximation is reasonable here.  Use the rule of thumb from the course notes.
::::

*****
Response

*****

:::: {.taskbox data-latex=""}
## Q3 (2 pts)

Use `prop.test( , correct = TRUE)` to calculate an (approximate) 95% confidence interval for $\pi$.  The resulting interval is generated using using Wilson’s score method.  You can just show the R output for full credit.
::::

```{r}
#Q3

```


:::: {.taskbox data-latex=""}
## Q4 (4 pts)

Using the confidence interval from the previous question, interpret the interval in context.  Aim for an interpretation that could be used in a write-up: avoid abbreviations and statistical jargon. 
::::

*****
Response

*****

:::: {.taskbox data-latex=""}
## Q5 (4 pts)

Construct an (approximate) 95% Wald confidence interval for $\pi$. Do this "by hand" (using R as a calculator) and echo your R code.
::::

***** 
Response

```{r, echo = TRUE}
#Q5

```

*****

:::: {.taskbox data-latex=""}
## Q6 (2 pts)

Use `binom.test()` to calculate the exact 95% confidence interval for $\pi$.  You can just show the R output for full credit.
::::

```{r}
#Q6

```

# Bomb Hits (Q7 - Q8)

This data is taken from Ott&Longnecker and originally from Hand (1993).  During WWII, bomb hits were recorded in the n = 576 grids in a map of a region of South London.  If the bomb hits were purely random, a Poisson model would produce the number of hits per grid.

| # Bomb Hits | 0   | 1   | 2  | 3  | 4+ | 
|-------------|---  |---  |--- |--- |----|
| # Grids     | 229 | 211 | 93 | 35 | 8  |

:::: {.taskbox data-latex=""}
## Q7 (2 pts)

Calculate $\hat\mu$, the sample mean hits per grid.  

Notes: (1) To do this calculation find the total number of bomb hits and divide by the number of grids. (2) For the calculation, treat 4+ as 4.  (3) Some example R code has been provided to get you started.  But other approaches are allowed.
::::

*****
Response

```{r}
#Q7
Bombs <- data.frame(Y = seq(from = 0, to = 4), 
                    Obs = c(229, 211, 93, 35, 8))

```

******

:::: {.taskbox data-latex=""}
## Q8 (8 pts)

Use the Pearson chi-square GOF test to test whether the number of bomb hits per grid follows the Poisson distribution.  Calculate the test statistic, df, p-value and give a conclusion (in context) using $\alpha$ = 0.05.  

Notes: (1) Do this "by hand" (using R as a calculator) but you can double check the test statistic using `chisq.test()`. (2) When calculating Poisson probabilities, you will use your $\hat\mu$ value from the previous question.  (3) When calculating Poisson probabilities, the last value should represent P($Y \ge 4$).  This can be done "manually".  (4) When calculating df, we "lose" 1 df because estimated $\hat\mu$.  
::::

*****
Response


```{r}
#Q8

```

*****


# Auto Accidents (Q9 - Q14)

CDA problem 2.4 concerns fatality results for drivers and passengers in auto accidents in Florida in 2008, according to whether the person was wearing a seat belt.

```{r}
#Q9-Q14
Auto <- matrix(c(441239, 703,
                  55623, 1085), nrow = 2, byrow = TRUE)
rownames(Auto) <- c("YesSeatBelt", "NoSeatBelt")
colnames(Auto) <- c("Nonfatal", "Fatal")
Auto
```

:::: {.taskbox data-latex=""}
## Q9 (3 pts)

Estimate (a) the marginal probability of **fatality**.  Also estimate the probability of fatality, conditional on seat-belt use in category (b) yes and (c) no.
::::

*****
Response

```{r}
#Q9

```

*****

:::: {.taskbox data-latex=""}
## Q10 (3 pt)

Estimate (a) the marginal probability of **wearing a seatbelt**.  Also estimate the probability of wearing a seatbelt, conditional on the injury being (b) Nonfatal and (c) Fatal.  
::::

*****
Response

```{r}
#Q10

```

*****

## Auto Accidents continued (Q11 - Q13)

Use injury as the response variable (with fatal injury being the "event of interest") and compare those not wearing seatbelts vs those wearing seatbelts.

:::: {.taskbox data-latex=""}
## Q11 (6 pts)

Provide an estimate of the **difference in proportions** and the corresponding (Wald) 95\% confidence interval.  Use the interval to make a conclusion about association between seat belt use and injury.  Be sure to mention the direction of association. (By direction of association, I just mean which group has the higher probability (risk) of fatal injury.)  
::::

*****
Response

```{r}
#Q11

```

*****

:::: {.taskbox data-latex=""}
## Q12 (6 pts)

Provide an estimate of the **relative risk** and the corresponding (Wald) 95\% confidence interval.   Use the interval to make a conclusion about association between seat belt use and injury.  Be sure to mention the direction of association.  (By direction of association, I just mean which group has the higher probability (risk) of fatal injury.)  
::::

*****
Response

```{r}
#Q12

```

*****

:::: {.taskbox data-latex=""}
## Q13 (6 pts)

Provide an estimate of the **odds ratio** and the corresponding (Wald) 95\% confidence interval.  Use the interval to make a conclusion about association between seat belt use and injury.  Be sure to mention the direction of association.  (By direction of association, I just mean which group has the higher odds of fatal injury.)  
::::

*****
Response

```{r}
#Q13

```

*****

:::: {.taskbox data-latex=""}
## Q14 (2 pts)

Considering the results of the previous two questions, why are the relative risk and odds ratio approximated equal?
::::

*****
Response

*****

# Appendix
```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}
```