---
title: "HW6_567"
author: "Matthew Stoebe"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

#Question 1
##a
Start by visualizing the Cauchy density using a red curve for −10 ≤ x ≤ +10. One option
is to use dcauchy(). For comparison, also overlay a normal density curve in blue.
```{r}

x <- seq(-10, 10, length.out = 1000)


cauchy_density <- dcauchy(x, location = 0, scale = 1)
normal_density <- dnorm(x, mean = 0, sd = 1)


plot(x, cauchy_density, type = "l", col = "red", lwd = 2,
     ylab = "Density", main = "Standard Cauchy and Normal Densities",
     xlab = "x")
lines(x, normal_density, col = "blue", lwd = 2)
legend("topright", legend = c("Cauchy", "Normal"),
       col = c("red", "blue"), lwd = 2)


```
##b
To prepare to use the inverse transform method, find F −1(u).
```{r}

set.seed(490)
n <- 100000

u <- runif(n, min = 0, max = 1)
x_cauchy <- tan(pi * (u - 0.5))


```

##c
(4 pts) Use the inverse transform method to generate a random sample of 10,000 observa-
tions from the standard Cauchy distribution. Show a histogram of your sample using a large
number of bins for your histogram (e.g., 100). In order to show all the data, do NOT restrict
your x axis. What striking features do you notice about your histogram?
```{r}
set.seed(123)

u <- runif(n, min = 0, max = 1)
x_cauchy <- tan(pi * (u - 0.5))

hist(x_cauchy, breaks = 100, main = "Histogram of Second Standard Cauchy Sample", xlab = "x")

```
There are some insane outliers every time this is run

##d
```{r}
# Set a different seed
set.seed(999)

u2 <- runif(n, min = 0, max = 1)
x_cauchy2 <- tan(pi * (u2 - 0.5))


hist(x_cauchy2, breaks = 100, main = "Histogram of Second Standard Cauchy Sample", xlab = "x")

```
More insane outliers which wont even show up on a plot

##e
The ratio of two independent standard normal (mean = 0, sd = 1) random variables is
distributed standard Cauchy. Use this information and rnorm() to generate a random sample
of 10,000 observations. Show the histogram.
```{r}


set.seed(789)

n <- 10000
Z1 <- rnorm(n, mean = 0, sd = 1)
Z2 <- rnorm(n, mean = 0, sd = 1)

# Compute the ratio
x_cauchy_ratio <- Z1 / Z2

# Plot histogram
hist(x_cauchy_ratio, breaks = 100, main = "Histogram of Standard Cauchy Sample (Ratio Method)",  xlab = "x")

```
Again, Because of the outliers this distribution is totally shifted from the one above. 

#Question 2

##a

```{r}
x_vals <- seq(-10, 10, length.out = 1000)


f_x <- exp(-x_vals) / (1 + exp(-x_vals))^2
g_x <- ifelse(x_vals >= 0, exp(-x_vals), exp(x_vals))


plot(x_vals, f_x, type = "l", col = "black", lwd = 2,
     xlab = "x", ylab = "Density",
     main = "Logistic Density f(x) and Exponential Density g(x)")

lines(x_vals, g_x, col = "blue", lwd = 2)


legend("topright", legend = c("Logistic Density f(x)", "Exponential Density g(x)"),
       col = c("black", "blue"), lwd = 2, bty = "n")

```


```{r}

# PARAMS
set.seed(420)
N <- 10000
samples <- numeric(N)
accepted <- 0 
total_trials <- 0

while (accepted < N) {
  x <- rexp(1, rate = 1)
  alpha <- (1 / (1 + exp(-x)))^2
  
  u <- runif(1)
  
  if (u <= alpha) {
    # Assign random sign
    s <- sample(c(-1, 1), 1)
    samples[accepted + 1] <- s * x
    accepted <- accepted + 1
  }
  total_trials <- total_trials + 1
}


acceptance_rate <- (N / total_trials) * 100
cat("Acceptance Rate:", round(acceptance_rate, 2), "%\n")


hist(samples, breaks = 50, probability = TRUE,
     main = "Histogram of Samples from Logistic Distribution",
     xlab = "Value", xlim = c(-10, 10), col = "lightblue", border = "white")

x_vals <- seq(-10, 10, length.out = 1000)
logistic_density <- exp(-x_vals) / (1 + exp(-x_vals))^2
lines(x_vals, logistic_density, col = "red", lwd = 2)

legend("topright", legend = c("Sample Histogram", "Logistic Density"),
       col = c("lightblue", "red"), lwd = c(10, 2), bty = "n")

```

