---
title: "Chapter 4 Homework Problems"
author: "Matthew Stoebe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Question 1
For the data in the table obtain the permutation distribution of the mean of the differences

```{r}
t1 <- c(100, 250, 50, 80)
t2<- c(112,240, 58, 82)


diff <- t1 - t2
diff_obs <- mean(diff)

# Number of permutations
n_perm <- 10000
perm_stats <- numeric(n_perm)

for (k in 1:n_perm){
  flip <- sample(c(1,-1), size = length(t1), replace = TRUE)
  perm_diff <- (t2 - t1) * flip
  
  perm_stats[k] <- mean(perm_diff)
}


cat("Percentage of Permutations with a more extreme difference than observed: ")
print(mean(perm_stats > diff_obs) * 100)

```

#Question 2
Students in an introductory statistics course were asked at the beginning of the course and at the end of the course the extent to which they disagreed or agreed with the statement "Statistics is important in my major area of study". They responded on a scale of 1 to 5 with 1 being stronly diagree and 5 being strongly agree. Data are shown in the table. Use a paired comparison permutation test to determine whether responses to this question changed significantly from the beginning to the end of the semester

```{r}
before <- c(2,3,4,4,3,1,3,4,4,5,3,4,2,2,4,3,4,2,2)
after  <- c(2,4,4,4,4,4,3,5,4,4,4,5,4,2,5,5,4,1,2)

diff_obs <- mean(after - before)
print(paste("Observed Mean Absolute Difference:", diff_obs))

# Number of permutations
n_perm <- 10000
perm_stats <- numeric(n_perm)

for (k in 1:n_perm){
  # For each pair, decide whether to flip the difference
  flip <- sample(c(1, -1), size = length(before), replace = TRUE)
  perm_diff <- (after - before) * flip
  
  # Calculate the mean absolute difference for this permutation
  perm_stats[k] <- mean(perm_diff)
}

# Calculate p-value
p_value <- mean(perm_stats >= diff_obs) * 100

cat("Percentage of Permutations with a more extreme difference than observed:", p_value, "%\n")

```

#Question 3
Obtain the permutation distribution of the signed rank statistic SR+ for the data in Exercise 1

```{r}
t1 <- c(100, 250, 50, 80)
t2<- c(112,240, 58, 82)


obs_diff_ranks <- rank(c(t2 - t1), ties.method="average")
obs_diff_signs <- sign(c(t2 - t1))

obs_signed_ranks <- obs_diff_ranks * obs_diff_signs
obs_sum_pos_ranks <- sum(obs_signed_ranks[obs_signed_ranks > 0])

cat("Observed sum of positive ranks:", obs_sum_pos_ranks, "\n")


n_perm <- 10000
perm_stats <- numeric(n_perm)

for (k in 1:n_perm){
  flip <- sample(c(1, -1), size = length(t1), replace = TRUE)
  signed_ranks <- rank(c(t2 - t1), ties.method="average") * flip
  perm_stats[k] <- sum(signed_ranks[signed_ranks > 0])
}

# Calculate p-value
p_value <- mean(perm_stats >= obs_sum_pos_ranks) * 100

cat("Percentage of Permutations with a more extreme difference than observed:", p_value, "%\n")

hist(perm_stats)
```

#Question 4
Measurements of the blood enzyme LDH were taken on seven subjects before fasting and after fasting. Is there a significant difference between the LDH readings before and after fasting? Test using Wilcox's signed-rank statistic. Compare the p-value of the exact test with that of the normal approximation.

```{r}
t1 <- c(89,90,87,98,120,85,97)
t2 <- c(76,101,84,86,105,84,93)

obs_diff_ranks <- rank(c(t2 - t1), ties.method="average")
obs_diff_signs <- sign(c(t2 - t1))

obs_signed_ranks <- obs_diff_ranks * obs_diff_signs
obs_sum_pos_ranks <- sum(obs_signed_ranks[obs_signed_ranks > 0])

cat("Observed sum of positive ranks:", obs_sum_pos_ranks, "\n")


n_perm <- 10000
perm_stats <- numeric(n_perm)

for (k in 1:n_perm){
  flip <- sample(c(1, -1), size = length(t1), replace = TRUE)
  signed_ranks <- rank(c(t2 - t1), ties.method="average") * flip
  perm_stats[k] <- sum(signed_ranks[signed_ranks > 0])
}

# Calculate p-value
p_value <- mean(perm_stats >= obs_sum_pos_ranks) * 100

cat("Percentage of Permutations with a more extreme difference than observed:", p_value, "%\n")

hist(perm_stats)

n = length(t1)

exp_sr = (n * (n+1))/4
var_sr = (n * (n+1) * (2*n + 1))/24

Z = (obs_sum_pos_ranks - exp_sr)/sqrt(var_sr)

pnorm(q=Z, lower.tail=FALSE)


```
#Question 5 6

To get the same process with ties, just copy above where i have method.ties = average

```{r}

```

