---
title: 'STAA 552: Final Exam Part 4'
author: "Your Name Here"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
#Retain this code chunk!!!
library(knitr)
library(tidyverse)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

**Honor Code from Part 1 applies here, too.**

# Epilepsy (Q20 - Q27)

To study the anti-epileptic drug progabide, researchers randomly assigned subjects suffering from epileptic seizures to receive either progabide or placebo.  The goal of the study is to evaluate whether progabide is effective in reducing seizures as compared to placebo.  We want to control for Age and the number or Pre-treatment seizures.  **Post is the response variable for all models.** We will work with n = 58 subjects.  The data is available from Canvas as epilepsy.csv.  

The data includes the following variables:

- ID: ID number (should NOT be used for model fitting)
- Age: age (in years)
- Trt: placebo or progabide
- Pre: number of epileptic seizures in the 8 weeks prior to start of treatment
- Post:  number epileptic seizures in the 8 weeks after start of treatment.  

## Q20 

Create a scatter plot showing number of seizures Post vs Pre treatment with observations color coded by Trt. Overlay separate Poisson regression curves (with log link) for each Trt.

```{r, out.width = "85%"}
#Q20
epilepsy <- read.csv("epilepsy.csv")
epilepsy$Trt <- factor(epilepsy$Trt, levels = c("placebo", "progabide"))


ggplot(epilepsy, aes(x=Pre, y=Post, color=Trt)) +
  geom_point() +
  stat_smooth(method="glm",
              method.args = list(family = poisson(link="log")),
              formula = y ~ x,
              se = FALSE) +
  theme_minimal()
```


## Q21

For **Model 1**, fit a Poisson regression model (with log link), using Post as the response and including Age, Trt and log(Pre) as predictors.   Show the coefficients table (including coefficient estimates and Wald test p-values).   

*****
```{r}
#Q21
Model1 <- glm(Post ~ Age + Trt + log(Pre), data = epilepsy, family = poisson)
summary(Model1)
```

*****

## Q22

Using **Model 1**, provide a detailed one sentence **interpretation** of the coefficient corresponding to **Trt** in context. 

*****
Response

```{r}
#Q22

exp_coef <- exp(coef(Model1)["Trtprogabide"])
exp_coef
```
Controlling for age and pre-treatment seizure frequency, patients on progabide have .78 times the post-treatment seizure rate of those on placebo

*****

## Q23 (2 pts)

Using **Model 1**, briefly discuss whether we should consider simplifying the model by dropping predictors.

*****
Response
Both features in model 1 are highly significant so we dont need to simplify the model
*****

## Q24 

Using **Model 1**, do we have evidence of lack of fit (or over-dispersion)?  Provide one piece of evidence and briefly discuss. (If multiple pieces of evidence are provided, the first will be graded.)

*****

```{r}
#Q24
pearson_res <- residuals(Model1, type = "pearson")

plot(fitted(Model1), pearson_res,
     xlab = "Fitted values", ylab = "Pearson Residuals",
     main = "Pearson Residuals vs Fitted Values")
abline(h = 0, col = "red", lty = 2)

prop_greater_2 <- sum(abs(pearson_res) > 2) / length(pearson_res)
cat("Proportion of residuals > 2 in absolute value:", prop_greater_2, "\n")
```

Discussion:
There are more extreme residuals than you would expect.

*****

## Q25 (2 pts)

Regardless of your answer to Q24, if we were concerned that there was lack of fit due to model "mis-specification", briefly mention one thing we could try (using available data).  

*****
Response
We would use a quasipoisson instead of regular poisson
*****

## Q26

Regardless of your answer to Q24, fit a model similar to Model 1 but allowing for over-dispersion.  Show the coefficients table (including coefficient estimates and Wald test p-values). We will call this **Model 2**.

*****
```{r}
#Q26
Model2 <- glm(Post ~ Age + Trt + log(Pre), data = epilepsy, family = quasipoisson)
summary(Model2)
```

*****

## Q27 (2 pts)

Based on **Model 2**, what can we conclude about effectiveness of Trt?

Note: Comparing the various epilepsy models, you may find that the conclusion about effectiveness of Trt varies. 

*****
Response
When we account for over dispersion in the data, we see that the effect is no longer significant. 
*****

# Appendix
```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}
```
