---
title: "565_HW_4"
author: "Matthew Stoebe"
date: "2025-04-15"
output: pdf_document
---

# Question 1

```{r}
library(dplyr)

site <- c(rep("A", 10), rep("B", 10), rep("C", 10), rep("D", 10))

defective <- c(57, 11, 57, 30, 119, 50, 31, 21, 67, 31,   # Site A
               39, 42, 133, 86, 3, 76, 41, 12, 114, 39,    # Site B
               28, 97, 34, 117, 6, 109, 33, 115, 65, 74,     # Site C
               72, 128, 73, 8, 74, 62, 78, 32, 8, 74)         # Site D
batch_size <- c(1000, 100, 1000, 500, 2000, 1000, 500, 500, 1000, 500,
                500, 500, 2000, 1000, 100, 1000, 500, 100, 2000, 500,
                500, 2000, 500, 2000, 100, 2000, 500, 2000, 1000, 1000,
                1000, 2000, 1000, 100, 1000, 1000, 1000, 500, 100, 1000)

chips_df <- data.frame(site, defective, batch_size)

agg_data <- chips_df %>%
  group_by(site) %>%
  summarize(total_def = sum(defective),
            total_chips = sum(batch_size),
            prop = total_def / total_chips)


x <- agg_data$total_def
n <- agg_data$total_chips

prop_test_result <- prop.test(x, n, correct = FALSE)
print(prop_test_result)

```

We see that there is a signifficant differences between the factories. the rate at factory 4 and factor 2 is higher than the rate at factory 1 and 3. this is found at a  alpha = .01 level
for this, we are assumign that each batch's defectsa re independent bernouli trials. 


# Question 2

Need to get the dataset set uop
```{r}


gpasat <- read.table("Data/gpasat.txt", header = FALSE)

colnames(gpasat) <- c("GPA", "SAT")

full_cor <- cor.test(gpasat$GPA, gpasat$SAT)
print(full_cor)

accepted     <- subset(gpasat, SAT + 200*GPA > 1550)
matriculated <- subset(accepted, SAT + 200*GPA <= 1850)

stu_cor <- cor.test(matriculated$GPA, matriculated$SAT)
print(stu_cor)

```

Conditioning on the admission rule forces a trade‑off—high SAT plus high GPA both aren’t needed to clear the cutoff—so within that band a higher score on one measure implies a lower score on the other, flipping the sign of the observed correlation





# Question 3

```{r}
beta0 <- 0.5
beta1 <- 2
beta2 <- 1.5
beta3 <- -1
sigma <- 0.5

x_val <- 1.2
d_clay <- 1
expected_height_clay <- beta0 + beta1 * x_val + beta2 * d_clay + beta3 * (x_val * d_clay)
cat("Expected height for clay soil with 1.2 gallons/week:", expected_height_clay, "inches\n")

x_val <- 1
expected_height_sandy <- beta0 + beta1 * x_val  # since d=0
expected_height_clay <- beta0 + beta1 * x_val + beta2 * 1 + beta3 * (x_val * 1)
cat("At 1 gallon/week:\n")
cat("  Sandy soil:", expected_height_sandy, "inches\n")
cat("  Clay soil:", expected_height_clay, "inches\n")


x_val <- 2
expected_height_sandy <- beta0 + beta1 * x_val  # d=0
expected_height_clay <- beta0 + beta1 * x_val + beta2 * 1 + beta3 * (x_val * 1)
cat("At 2 gallons/week:\n")
cat("  Sandy soil:", expected_height_sandy, "inches\n")
cat("  Clay soil:", expected_height_clay, "inches\n")


mu <- beta0 + beta1 * 1 + beta2 * 1 + beta3 * (1 * 1)  # mean = 3.0 inches
threshold <- 4

p_above4 <- 1 - pnorm(threshold, mean = mu, sd = sigma)
cat("Probability that grass in clay soil with 1 gallon/week is taller than 4 inches:", p_above4, "\n")

```


# Question 4

\noindent \textbf{(a) Notation \& Model:} \\
Let \(Y_i\) be the dry weight of the \(i\)th seedling, and let \(x_i\) be the dose (in ounces) applied. Define 
\[
D_i = \begin{cases}
0, & \text{if seedling } i \text{ receives booster A},\\[5mm]
1, & \text{if seedling } i \text{ receives booster B}.
\end{cases}
\]
For the 12 treated seedlings (ignoring controls), a suitable linear model is
\[
Y_i = \beta_0 + \beta_1 x_i + \beta_2 D_i + \beta_3 (x_i \cdot D_i) + \varepsilon_i, \quad \varepsilon_i \sim N(0,\sigma^2).
\]

\noindent \textbf{(b) Parameter Interpretation:} \\
\(\beta_0\): Baseline dry weight for booster A (when \(x_i = 0\)).\\[4mm]
\(\beta_1\): Increase in dry weight per ounce of dose for booster A.\\[4mm]
\(\beta_2\): Difference in baseline weight (at \(x_i = 0\)) between booster B and A.\\[4mm]
\(\beta_3\): Difference in the dose effect between boosters; so for booster B the dose-response slope is \(\beta_1+\beta_3\).


\noindent \textbf{(c) Hypotheses:} \\
To test if the two boosters are equally effective (i.e., have the same baseline and dose effect), we set
\[
\begin{aligned}
H_0: \quad & \beta_2 = 0 \text{ and } \beta_3 = 0, \\
H_a: \quad & \text{at least one of } \beta_2 \text{ or } \beta_3 \text{ differs from } 0.
\end{aligned}
\]


