---
title: "551 Assignment 2"
author: "Matthew Stoebe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework Questions:

##Chapter 4: 
###4.10

Survey weighting: Compare two options for a national opinion survey: (a) a simple random
sample of 1000 Americans, or (b) a survey that oversamples Latinos, with 300 randomly sampled
Latinos and 700 others randomly sampled from the non-Latino population. One of these options
will give more accurate comparisons between Latinos and others; the other will give more
accurate estimates for the total population average.

(a) Which option gives more accurate comparisons and which option gives more accurate
population estimates?

(b) Explain your answer above by computing standard errors for the Latino/other comparison and
the national average under each design. Assume that the national population is 15% Latino,
that the items of interest are yes/no questions with approximately equal proportions of each
response, and (unrealistically) that the surveys have no problems with nonresponse.
```{r}

```

##Chapter 5: 
###5.2, 

Continuous probability simulation: The logarithms of weights (in pounds) of men in the United
States are approximately normally distributed with mean 5.13 and standard deviation 0.17;
women’s log weights are approximately normally distributed with mean 4.96 and standard
deviation 0.20. Suppose 10 adults selected at random step on an elevator with a capacity of 1750
pounds. What is the probability that their total weight exceeds this limit?


```{r}

```

###5.6, 
Propagation of uncertainty: We use a highly idealized setting to illustrate the use of simulations
in combining uncertainties. Suppose a company changes its technology for widget production,
and a study estimates the cost savings at $5 per unit, but with a standard error of $4. Furthermore,
a forecast estimates the size of the market (that is, the number of widgets that will be sold)
at 40 000, with a standard error of 10 000. Assuming these two sources of uncertainty are
independent, use simulation to estimate the total amount of money saved by the new product
(that is, savings per unit, multiplied by size of the market).

```{r}

```

###5.10
Inference for a ratio of parameters: A (hypothetical) study compares the costs and effectiveness
of two different medical treatments.
• In the first part of the study, the difference in costs between treatments A and B is estimated
at $600 per patient, with a standard error of $400, based on a regression with 50 degrees of
freedom.
• In the second part of the study, the difference in effectiveness is estimated at 3.0 (on some
relevant measure), with a standard error of 1.0, based on a regression with 100 degrees of
freedom.
• For simplicity, assume that the data from the two parts of the study were collected independently.
Inference is desired for the incremental cost-effectiveness ratio: the difference between the
average costs of the two treatments, divided by the difference between their average effectiveness,
a problem discussed further by Heitjan, Moskowitz, and Whang (1999).
(a) Create 1000 simulation draws of the cost difference and the effectiveness difference, and make
a scatterplot of these draws.
(b) Use simulation to come up with an estimate, 50% interval, and 95% interval for the incremental
cost-effectiveness ratio.
(c) Repeat, changing the standard error on the difference in effectiveness to 2.0.
```{r}

```

##Chapter 6: 
###6.2, 
Programming fake-data simulation: Write an R function to: (i) simulate n data points from
the model, y = a + bx + error, with data points x uniformly sampled from the range (0, 100)
and with errors drawn independently from the normal distribution with mean 0 and standard
deviation σ; (ii) fit a linear regression to the simulated data; and (iii) make a scatterplot of the
data and fitted regression line. Your function should take as arguments, a, b, n, σ, and it should
return the data, print out the fitted regression, and make the plot. Check your function by trying
it out on some values of a, b, n, σ.

```{r}

```

###6.3
Variation, uncertainty, and sample size: Repeat the example in Section 6.2, varying the number
of data points, n. What happens to the parameter estimates and uncertainties when you increase
the number of observations?
```{r}

```

##Chapter 7: 

###7.2,
Fake-data simulation and regression: Simulate 100 data points from the linear model, y =
a + bx + error, with a = 5, b = 7, the values of x being sampled at random from a uniform
distribution on the range [0, 50], and errors that are normally distributed with mean 0 and standard
deviation 3.
(a) Fit a regression line to these data and display the output.
(b) Graph a scatterplot of the data and the regression line.
(c) Use the text function in R to add the formula of the fitted line to the graph.
```{r}

```

###7.6
Formulating comparisons as regression models: Take the election forecasting model and simplify
it by creating a binary predictor defined as x = 0 if income growth is less than 2% and x = 1 if
income growth is more than 2%.
(a) Compute the difference in incumbent party’s vote share on average, comparing those two
groups of elections, and determine the standard error for this difference.
(b) Regress incumbent party’s vote share on the binary predictor of income growth and check
that the resulting estimate and standard error are the same as above.

```{r}

```
