---
title: "STAA 551 Assignment 1"
author: "Matthew Stoebe"
date: "`r Sys.Date()`"
output: word_document
---

# Assignment Description
Use R Markdown to complete your assignment.

All problem numbers refer to Regression and Other Stories by Gelman, Hill, and Vehtari
Submission is a PDF file upload. 
There is only one attempt for this assignment, so make sure you upload the correct file

- Chapter 2: 2.3, 2.9
- Chapter 3: 3.6--3.9

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

install.packages(c(
  "dplyr", 
  "tidyr", 
  "ggplot2"
))

library(dplyr)
library(tidyr)
library(ggplot2)

```

# Question 2.3

```{r pressure, echo=FALSE}

allnames <- read.csv("./ROS-Examples-master/Names/data/allnames_clean.csv")

female_data <- filter(allnames, sex == "F")

female_data$last_letter <- substr(female_data$name, nchar(female_data$name), nchar(female_data$name))

grouped_data <- group_by(female_data, last_letter)


summarized_data <- summarise(grouped_data, across(matches("^X\\d"), sum, na.rm = TRUE))


total_sums <- colSums(summarized_data[, -1])  # Exclude the first column (last_letter)

percentage_data <- summarized_data

for (col in names(summarized_data[-1])) {
  percentage_data[[col]] <- 100 * (percentage_data[[col]] / total_sums[col])
}

percentage_long <- pivot_longer(percentage_data, cols = starts_with("X"), 
                                names_to = "year", 
                                values_to = "percentage")
percentage_long <- percentage_long %>%
  mutate(year = as.numeric(sub("^X", "", year)))


top_labels_2010 <- percentage_long %>%
  filter(year == 2010) %>%
  arrange(desc(percentage)) %>%
  slice_head(n = 5)  # Get the top 3 lines based on percentage

top_labels_1910 <- percentage_long %>%
  filter(year == 1910) %>%
  arrange(desc(percentage)) %>%
  slice_head(n = 5)  # Get the top 3 lines based on percentage



ggplot(percentage_long, aes(x = year, y = percentage, color = last_letter, group = last_letter)) +
  geom_line(size = 1) +
  labs(title = "Last Letters of girl's names",
       x = "Year",
       y = "Percentage",
       color = "Last Letter") +
  scale_x_continuous(breaks = seq(1880, 2010, by = 10)) +  # Show labels for every 10 years
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +

  geom_text(data = top_labels_2010,
            aes(label = last_letter),
            vjust = -0.5, hjust = 1.5, color = "black") +
  
  # Optional: Add custom annotations or vertical lines as needed
  geom_vline(xintercept = 2010, linetype = "dashed", color = "black") +

  geom_text(data = top_labels_1910,
            aes(label = last_letter),
            vjust = -0.5, hjust = 1.5, color = "black") +
  
  # Optional: Add custom annotations or vertical lines as needed
  geom_vline(xintercept = 1910, linetype = "dashed", color = "black")

```



# Question 2.9

Graphing parallel time series: The mortality data in Section 2.4 are accessible from this site at
the U.S. Centers for Disease Control and Prevention: wonder.cdc.gov. Download mortality data
from this source but choose just one particular cause of death, and then make graphs similar to
those in Section 2.4, breaking down trends in death rate by age, sex, and region of the country

```{r}
data_path <- "./Other Data/Mortality/Underlying_Cause_of_Death_2018_2022_Single_Race.csv"

# Initial Data Prep
#data <- read.delim("C:\\Users\\mstoebe\\Downloads\\Underlying Cause of Death, 2018-2022, Single Race (1).txt", sep = "\t", header = TRUE)
#filtered_data <- data[!is.na(data$`Year.Code`) & data$`Year.Code` != "", ]
#write.csv(filtered_data, data_path)

data <- read.csv(data_path)

data$Crude.Rate <- as.numeric(data$Crude.Rate)

filtered_data <- data[!is.na(data$Crude.Rate), ]

# Convert Population and Deaths to numeric, setting non-numeric values to NA
filtered_data$Population <- as.numeric(gsub(",", "", filtered_data$Population))
filtered_data$Deaths <- as.numeric(gsub(",", "", filtered_data$Deaths))
filtered_data$Single.Year.Ages <- as.numeric(gsub("[^0-9]", "", filtered_data$`Single.Year.Ages`))
filtered_data$Age.Bracket <- floor(filtered_data$Single.Year.Ages / 10) * 10
filtered_data$Age.Bracket <- factor(filtered_data$Age.Bracket, 
                                    levels = seq(0, 90, by = 10), 
                                    labels = paste(seq(0, 90, by = 10), seq(9, 99, by = 10), sep = "-"))


summary_by_year <- aggregate(cbind(Deaths, Population) ~ Year, data = filtered_data, sum, na.rm = TRUE)
summary_by_year$Crude.Rate <- (summary_by_year$Deaths / summary_by_year$Population) * 100000


ggplot(summary_by_year, aes_string(x = "Year", y = "Crude.Rate")) +
  geom_line() + 
  labs(title = "Crude Death Rate for Tummy Aches by Year",
       x = "Year",
       y = "Crude Death Rate per 100,000") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

features <- c("Age.Bracket","Gender.Code", "Single.Race.6", "Census.Region.Code" )


for (feature in features) {
  formula <- as.formula(paste("cbind(Deaths, Population) ~ Year +", feature))
  summary_by_feature <- aggregate(formula, data = filtered_data, sum, na.rm = TRUE)
  summary_by_feature$Crude.Rate <- (summary_by_feature$Deaths / summary_by_feature$Population) * 100000

  p <- ggplot(summary_by_feature, aes_string(x = "Year", y = "Crude.Rate", color = feature)) +
    geom_line() +
    geom_point() +
    labs(title = paste("Crude Death Rate by Year and", feature),
         x = "Year",
         y = "Crude Death Rate per 100,000",
         color = feature) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))

  
  # Print the plot
  print(p)  

}


average_age_by_year <- aggregate(Single.Year.Ages ~ Year, data = filtered_data, mean, na.rm = TRUE)
ggplot(average_age_by_year, aes(x = Year, y = Single.Year.Ages)) +
  geom_line() + 
  geom_point() +
  labs(title = "Average Age by Year",
       x = "Year",
       y = "Average Age") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

