---
title: "572_Chapter5_Hw"
author: "Matthew Stoebe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Question 3
The data in the table are the ages in days of concrete cylinders and the compression strengths of the cylinders. 
a. Plot the data to show a nonlinear relationship. Compute Pearsons, Spearmans and Kendall's Tau
b. Test for significant association using each of the measures of association in part a

```{r pressure, echo=FALSE}
Age <- c(3,7,15,24,85,180,360)
Strength <- c(2500,3200,4300,5300,5900,6700,6900)

plot(Age, Strength, 
     main = "Age vs. Compression Strength of Concrete Cylinders",
     xlab = "Age (days)", 
     ylab = "Compression Strength (psi)",
     pch = 19,        # Solid circle
     col = "blue")    # Color of the points

lines(smooth.spline(Age, Strength), col = "red", lwd = 2)

print("Automatic calculations")
# Pearson's Correlation
pearson_cor <- cor(Age, Strength, method = "pearson")
print(paste("Pearson's r:", round(pearson_cor, 4)))

# Spearman's Rank Correlation
spearman_cor <- cor(Age, Strength, method = "spearman")
print(paste("Spearman's rho:", round(spearman_cor, 4)))

# Kendall's Tau
kendall_cor <- cor(Age, Strength, method = "kendall")
print(paste("Kendall's Tau:", round(kendall_cor, 4)))

```
```{r}
print("Manual calculations")

n<- length(Age)

# Step 1: Calculate Means
mean_Age <- sum(Age) / n
mean_Strength <- sum(Strength) / n

# Step 2: Calculate Deviations from Means
dev_Age <- Age - mean_Age
dev_Strength <- Strength - mean_Strength

# Step 3: Calculate Numerator (Covariance)
numerator <- sum(dev_Age * dev_Strength)

# Step 4: Calculate Denominator (Product of Standard Deviations)
denominator <- sqrt(sum(dev_Age^2) * sum(dev_Strength^2))

# Step 5: Calculate Pearson's r
pearson_r <- numerator / denominator

# Display the result
cat("Manual Calculation of Pearson's r:", round(pearson_r, 4), "\n")

rank_Age <- rank(Age, ties.method = "average")
rank_Strength <- rank(Strength, ties.method = "average")

# Step 2: Calculate Differences in Ranks
d <- rank_Age - rank_Strength

# Step 3: Calculate d^2
d_squared <- d^2

# Step 4: Sum of d^2
sum_d_squared <- sum(d_squared)

# Step 5: Calculate Spearman's rho
spearman_rho <- 1 - (6 * sum_d_squared) / (n * (n^2 - 1))

# Display the result
cat("Manual Calculation of Spearman's rho:", round(spearman_rho, 4), "\n")

calculate_kendall_tau <- function(x, y) {
  C <- 0  # Concordant pairs
  D <- 0  # Discordant pairs
  n <- length(x)
  
  for(i in 1:(n-1)) {
    for(j in (i+1):n) {
      # Compare pairs
      if((x[i] < x[j] && y[i] < y[j]) || (x[i] > x[j] && y[i] > y[j])) {
        C <- C + 1
      } else if((x[i] < x[j] && y[i] > y[j]) || (x[i] > x[j] && y[i] < y[j])) {
        D <- D + 1
      }
      # Ties are not considered in this basic implementation
    }
  }
  
  tau <- (C - D) / (0.5 * n * (n - 1))
  return(tau)
}

# Calculate Kendall's Tau
kendall_tau <- calculate_kendall_tau(Age, Strength)

# Display the result
cat("Manual Calculation of Kendall's Tau:", round(kendall_tau, 4), "\n")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
