---
title: "STAA 553: HW1"
author: "YOUR NAME HERE"
output: pdf_document
---

```{r setup, include=FALSE}
#Retain this code chunk!!!
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

See Canvas Calendar for due date.  
14 points total, 2 points per problem unless otherwise noted.  
Add or delete code chunks as needed.  
Content for all questions is from Sections 1 and 2.  


# Hand Washing 1 (Q1 - Q5)

An investigator is planning a hand-washing study.  They want to evaluate the effect of water temperature (60, 80, 100 or 120 F) on bacterial count on people’s palms after hand-washing.  They plan to recruit a total of n = 32 subjects and will randomly assign each subject to wash their hands with a single water temperature.

## Q1

Identify the experimental units.

*****
Response 

*****

## Q2

Identify the treatment (or factor) and number of levels. 

*****
Response

*****

## Q3

Identify the response variable.

*****
Response

*****

## Q4

Suggest one approach that could be used to "reduce noise" when conducting this study.  This is a "common sense" question with many possible correct answers, not something you will find in a textbook.

*****
Response

*****

## Q5

Use R `sample()` to randomly assign 32 subjects to temperatures (60, 80, 100 or 120 F) requiring balance (equal number of subjects per temperature).  Show a summary table giving the number of subjects per treatment.

*****
```{r}
#Q5

```
*****

# Hand Washing 2 (Q6 - Q7)

We continue with the hand-washing study.  But now suppose that each subject will be asked to wash their hands 4 times (on 4 different days), such that each subject experiences all 4 water temperatures.  This is an example of a blocked or repeated measures design. 

## Q6

Name one benefit of this design as compared to the original design.

*****
Response

*****

## Q7

Suggest one way that randomization could be incorporated into this study design.

*****
Response

*****

# Appendix
```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}
```