---
title: "STAA 553: HW6"
author: "YOUR NAME HERE"
output: pdf_document
---

```{r setup, include=FALSE}
#Retain this code chunk!!!
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```
 
See Canvas Calendar for due date.  
40 points total, 2 points per problem unless otherwise noted.  
Content for all questions is from section 10 or earlier.  

**Important Notes**:  
- Remember to check `str()` and then define things `as.factor` where needed.  
- A key feature of the balanced designs considered in this assignment are that the SE's are the same for all means and pairwise comparisons.

# Cell Line Study (Q1 - Q11)

DAEWR gives data (originally from Lew (2007)) from an experiment to determine whether cultured cells respond to two drugs. The study was conducted as a series of six experimental "runs" using a stable cell line plated onto Petri dishes.  Each experimental run included three Petri dishes: one treated with drug 1, one treated with drug 2, and one untreated serving as a control.  Hence, experiment (Exp) serves as a blocking variable.  The data is available from Canvas as “CellData.csv”.  
**See important notes above!**

## Q1

Provide a visual summary of the data using code similar to what is provided.

```{r, out.width = "70%"}
library(tidyverse)
#Q1
#ggplot(aes(x = Drug, y = Response, group = Exp, color = Exp), data = CellData) + 
#  geom_point() +
#  geom_line()
```

# Cell Line RCB1 (Q2 - Q5)

Analyze the data as an RCB design with **fixed blocks** using `lm()`.

## Q2

Show the Type 3 ANOVA table.

*****
```{r}
#Q2

```

*****

## Q3

Using the ANOVA table from the previous question, briefly discuss the effectiveness of the blocking for this study.   

*****
Response

*****

## Q4

Run Tukey adjusted pairwise comparisons for Drug and show the results.

*****
```{r}
#Q4

```

*****

## Q5

Show the residual diagnostic plots (Resids vs Fitted and QQplot of Resids).

```{r, out.width = "70%"}
#Q5

```

# Cell Line RCB2 (Q6 - Q9)

Analyze the data as an RCB design with **random blocks** using `lme4::lmer()`.

## Q6 (4 pts)

Provide estimates of the block variance and the residual variance. 

*****
$\sigma^2_{Block} = \sigma^2_{Exp} = XXXX$  
$\sigma^2_{\epsilon} = \sigma^2_{Resid} = XXXX$  

```{r}
#Q6

```

*****

## Q7

Show the Type 3 ANOVA table.  Notes: (1) In the mixed model framework, we can use `anova()` to do this after loading the lmerTest and pbkrtest packages. (2) The F-test for Drug should exactly match your result from Q2.

*****
```{r}
#Q7

```

*****

## Q8

Run Tukey adjusted pairwise comparisons for Drug and show the results. Note: These results should exactly match the pairwise comparisons from Q4.

*****
```{r}
#Q8

```

*****

## Q9

Show the residual diagnostic plot (Resids vs Fitted).

```{r, out.width = "70%"}
#Q9

```

# Cell Line CRD (Q10-Q11)

Just for illustration, analyze the data using a one-way model **ignoring blocks**.

## Q10

Show the ANOVA table.

*****
```{r}
#Q10

```
*****

## Q11

Compare the ANOVA tables from the RCB model (Q2 or Q7) vs the one-way/CRD model (Q10).  Briefly explain why we were able to detect differences between Drugs using the RCB model, when we did not detect differences using the one-way/CRD model. Your answer should be based on *specific output*.  Hint: You may want to calculate MSResid.

*****
Response

*****

# Sleep Study (Q12 - Q16)

DAEWR gives data (originally from le Riche and Csima (1964) from an experiment to evaluate 5 drug treatments (A, B, C, D, E) and their effect on sleep quality in elderly patients.  Higher values indicate better sleep quality. A latin square design was used to account for patient-to-patient and week-to-week differences.   The data is available from Canvas as "SleepData.csv".  A visual summary of the design (and results) is also available from Canvas.  
**See important notes above!**

## Q12 (4 pts)

Fit an appropriate fixed effects model using `lm()`. Show the Type 3 ANOVA table. 

*****
```{r}
#Q12

```

*****

## Q13

Using the ANOVA table from the previous question, briefly discuss the effectiveness of the latin square design for this study.  

*****
Response

*****

## Q14

Run Tukey adjusted pairwise comparisons for Trt and show the results.

*****
```{r}
#Q14

```

***** 

### Q15

Considering your results from the previous question, briefly summarize your findings.  

## Q15A

Discuss comparisons of active treatments (B, C, D, E) vs placebo (A) in context. 

*****
Response

*****

### Q15B

Discuss comparisons between the active treatments (B, C, D, E) in context. 

*****
Response

*****

## Q16 (4 pts)

Now re-analyze the data as a mixed model using `lme4::lmer()`. Use a model similar to Q12, but decide which terms should be included as fixed and random effects. Show the Type 3 ANOVA table.  Note: Due to balance, the F-test for Trt should exactly match your result from Q12.

*****
```{r}
#Q16

```
*****


# Appendix
```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}
```