---
title: "STAA 553: Final Exam"
subtitle: "Spring 2025"
output: pdf_document
---

```{r setup, include=FALSE}
#Retain this code chunk!!!
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

**Honor Pledge:** I have worked independently on this exam. I have read the exam instructions. I have not given, received, or used any unauthorized assistance on this exam.   
**YOUR NAME HERE**

**Instructions:** 

- This exam is due by Saturday 3/15 at midnight. 
- **Students are required to work independently on the exam.**  Do NOT discuss the exam with anyone else (including other students).
- You may use the textbook, class notes, examples, HW solutions posted in the current Canvas course.  You may use any other publicly available (print or online) statistics references or resources that you find helpful.  Use of homework “helper” websites (ex: Chegg, NoteHall, etc) is NOT allowed.  Use of chatbots (ex: ChatGPT) is NOT allowed. 
- Knit frequently to avoid last minute problems.  You may add or delete code chunks as needed.  **It is the student’s responsibility to check the knitted document (for correctness and completeness) before submitting.**
- For any questions that require calculations, you should provide R code for full credit.
- For some questions, there may be more than one possible answer, analysis or graph that could be used for full credit. **Choose one approach**, making a reasonable choice and justifying if needed. 
- Given this is the final exam, you should present your best work. I will deduct points for things like printing full data to knitted document, unreadable tables, unclear, excess or unnecessary output, etc.
- Use $\alpha$ = 0.05 and/or 95\% confidence where needed.
- All questions are worth 4 points except where noted.  Maximum score is 100.
- **I believe all students can do well on this exam.  Please don’t cheat!!!**




\newpage

# Blueberries (Q1 - Q4)

A study is being **planned** to compare 3 fertilizer treatments (A, B, C) on blueberry bushes. The response variable is yield (in lbs) **per bush** at the end of the growing season. The study will be run at a single farm, using existing mature blueberry bushes.   At this farm, there are a large number of fields of blueberry bushes representing different varieties and ages.  A single field has multiple bushes of a single age and single variety.   Initial soil nitrogen varies across locations (even within a field).  It is known that yield will be associated with variety, age and initial soil nitrogen.  

As a statistician, you are asked to make recommendations to design the study.  

## Q1 (2 pts)

Discuss how **replication** would be incorporated into the design. In other words, what is a replicate?

*****
Response 

*****

## Q2 (2 pts)

Discuss how **randomization** would be incorporated into the design. 

*****
Response

*****


## Q3A 

Blocking is one approach that can be used to **reduce noise**.  Provide a brief but specific description of how **blocking** could be incorporated into this research scenario.  How will you define/select blocks?  How will treatments be applied? 

*****
Response

*****

## Q3B 

Propose one other approach that can be used to **reduce noise**.  Since you already discussed blocking in the previous question, discuss something **different** here. (This can be an alternative to blocking or it may be used with blocking.)  The proposed method does not have to be complicated.  Provide a reasonable amount of detail specific to this research scenario and providing guidance about how to implement the method (1-3 sentences).  

*****
Response  

*****


## Q4 

Regardless of your answers to previous question, suppose the investigators plan to use a completely randomized design. They conjecture that $\mu_1$ = 32, $\mu_2$ = 34, $\mu_3$ = 40 and $\sigma$ = 8. Find the sample size required (number of blueberry bushes per treatment) to achieve 80\% power. Remember to state your answer as an integer value.

*****
n = 

```{r}
#Q4

```
*****

# Sheep (Q5 - Q8)

An investigator is interested in comparing 6 treatments (Trt 1-6) for bone injuries in sheep. To evaluate the treatments, he creates (and then treats) small bone defects in the femurs of sheep. The response variable (Y) is defect size at one month post treatment. They will use a total of 18 sheep (blocks) with 4 defects per sheep (block). They are considering a **balanced incomplete block design (BIBD)**.

## Q5

Identify the values for t (# treatments), r (# replicates per treatment), b (# blocks) and k (# "units" per block).  

*****

t =    

r =     

b =    

k =    

```{r}
#Q5

```

*****

## Q6 (6 pts)

**State and check** the 3 conditions for a BIBD. Using your values from the description and responses to the previous question, is a BIBD possible?  
Note: Even if there are mistakes in your responses to Q5, I am very open to giving partial credit as long as you show your work here.  

*****
Condition1:      

Condition2: 

Condition3:   

Is a BIBD possible?    

```{r}
#Q6

```

*****

## Q7

Regardless of your answer to the previous question, suppose the study will be run using a blocked design, even if we cannot achieve a balanced design.

### Q7A (4 pts)

If the design is imbalanced, how does this affect the **analysis**? In other words, are changes required to the model or to generate the ANOVA table depending on whether we have an IBD (incomplete block design) versus a BIBD?

*****
Response

*****

### Q7B (4 pts)

If the design is imbalanced, how does this affect the results? In other words, what  benefit of the BIBD is lost if we have an IBD?  Your discussion should be brief but still be specific. 

*****
Response

*****

### Q8 

The investigators plan to use Tukey's method to run pairwise comparisons between the 6 treatments.

### Q8A (4 pts)

Discuss the primary **benefit** of using Tukey's method (as compared to unadjusted pairwise comparisons). Your discussion should be brief but still be specific.     

*****
Response

*****

### Q8B (4 pts)

Discuss the primary **limitation** of using Tukey's method (as compared to unadjusted pairwise comparisons).     

*****
Response

*****

# Drug Study (Q9 - Q11)

A study was done to compare three drug Formulations (A = 5mg tablet, B = 100mg tablet, C = sustained release capsule). A total of n = 12 healthy adults volunteered to participate in the study. The study includes three periods, during which each subject received one of the formulations. There was a week long washout between each period. Subjects received the formulations in one of three sequences: ABC, BCA, CAB with 4 subjects randomly assigned to each sequence.

The data includes 36 rows and these variables:

- Subject: 1-12
- Formulation: A,B,C
- Period: 1, 2, 3
- Sequence: ABC, BCA, CAB
- PostBP: response, blood pressure after treatment

Notes:

-  No data is provided! You have to think about the analysis based on the description above.
- Not all variables may be needed for analysis.

## Q9 (2 pts)

Identify the ("blocking") design.  Just provide the name of the design, no need to justify. 

*****
Response   

*****

## Q10 

Considering your answer to the previous question, provide R code to fit an appropriate model. Note: I am primarily looking for a single line of code to set up the model. If you feel additional lines of code are needed, that is fine.

*****
Response

*****

## Q11

It is well known that blood pressure is variable even within a single person. Just for this question, suppose we also recorded the PreBP (blood pressure before treatment) for each subject and period.

### Q11A (2 pts)

Suggest one way to incorporate this information into the analysis.

*****
Response 

*****

### Q11A (2 pts)

Modify your code from Q10 to incorporate your suggestion.

*****
Response  

*****

# Pine Beetles (Q12 - Q16)

Treatments for pine beetle infestation are suspected of affecting the “hardness” of the wood. A researcher is interested in comparing the wood hardness (Y) for three treatments (Trt: A, B, C). The researcher begins the experiment by identifying 5 infested trees (Tree: 1-5). For each tree, six branches are randomly assigned to treatments with each Trt represented exactly twice for each tree. At the end of the study, the hardness is measured for each branch. Hence, we have 5 trees x 3 treatments x 2 branches = 30 observations. The ANOVA table (from Anova with type = 3 and appropriate contrasts) is given here.

| Source    | df |   SS | F value | p-value |
|-----------|---:|-----:|---------|---------|
| Tree      | 1  |   70 | 0.7303  | 0.4006  |
| Trt       | 2  |   38 | 0.1984  | 0.8213  |
| Residuals | 26 | 2493 |         |         |


## Q12 

Briefly identify the blocking and treatment design for this study. Include identifying details about the blocks (if applicable) and treatments.  

*****

Blocking Design:  

Treatment Design:  

*****

## Q13 (2 pts)

The Type 3 ANOVA table is shown above. Would the Type 1 ANOVA SS and F-tests have been the same or different? Just answer same or different. No need to justify.

*****
Response

*****

## Q14 (2 pts)

The analyst made a mistake when setting up the model. Identify the mistake. Note: This question is NOT looking for a discussion of whether Tree should be fixed or random.

*****
Response

*****

## Q15

After addressing the mistake from Q14, a colleague expresses concern that the assumption of independence is “not satisfied because you have repeat observations on trees”. Briefly discuss whether you agree with the colleague’s concern. If so, propose a modelling approach that can be used to address this concern.  

*****
Response  

*****

## Q16

After addressing the mistake from Q14, could an interaction (Tree:Trt) have been included in the model? Briefly discuss.

*****
Response

*****

# Feed Trial (Q17 - Q23)

A study was done to investigate a new vitamin supplement for cattle.  A successful supplement would increase **ADG** (average daily gain) in cattle. Two levels of **Suppl** (no or yes) and four **Feed**  formulations (A, B, C, D) and  were considered. The primary research question is about the effect of Suppl.  But the researchers are concerned that the effect of supplement could vary across feed types. A total of 24 **Pen**s of cattle were randomly assigned to treatment combinations such that there are exactly 3 pens per treatment combination. ADG was recorded (for each pen) at the end of the study. The data is available from Canvas as FeedTrial.csv. Use alpha = 0.05.

## Q17

Briefly identify the blocking and treatment design for this study. Include identifying details about the blocks (if applicable) and treatments.  

*****
Blocking Design:   

Treatment Design:    

*****

## Q18 (6 pts)

Fit an appropriate model and show an appropriate ANOVA table in your exam.

*****
```{r}
#Q18

```

*****

## Q19

Create an appropriate summary plot of the data.

*****
```{r, out.width = "70%"}
#Q19

```

*****

## Q20 (6 pts)

Provide two (common) diagnostic plots. Briefly discuss each plot and whether assumptions are satisfied. Notes: (1) Your discussion should be brief but still be specific. (2) Depending on how you set up your model, you may get a red note “hat values (leverages) are all = XXXX and there are no factor predictors”. This message can safely be ignored.

*****
Response

```{r}
#Q20

```

*****

## Q21

Considering the primary research question and your ANOVA table, provide appropriate pairwise comparisons. Use Tukey adjustment if appropriate.

*****

```{r}
#Q21

```

*****

## Q22

Briefly summarize your conclusions from the previous question, being sure to address the research question. Avoid statistical jargon (reject, hypotheses, etc).

*****
Response 

*****

## Q23

Just for this question, consider the output from code like the following:

```{r, include = FALSE}
#Q23
#emmeans(FeedModel, pairwise ~ Suppl)
#emmeans(FeedModel, pairwise ~ Feed)
```

Specifically, consider the contrasts “no – yes” vs “B – C”. Surprisingly, the magnitude of the estimated difference is larger for “B-C” but the p-value is also larger for “B-C”. This indicates lower power despite the larger magnitude difference. Give **two different** reasons why this happened. Note: You do not need show the output, just discuss the result

*****
Reason 1:  

Reason 2:    

*****


# Appendix
```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}
```