---
title: "STAA_551_Case_Study_ReneeMyers"
author: "Renee Myers"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
---
Variable Notes:
Golfer: Name of golfer, can exclude from the linear regression analysis
rounds: Number of rounds played, continuous variable. 
avedist: Average drive distance. Drive distance is affected by both age and gender, ranges between 243 yards and 196 yards on average. 
pctfrwy: Percent of fairways hit. Measures the number of times a golfer's tee shot lands in the fairway compared to the number of holes played. Gives an indication of player accuracy, most professionals hit 60-70% of fairways. 
pctgrn: Percentage of greens in regulation. Measurement of whether the ball finishes on the green in two strokes fewer than the par of the hole. Hitting a green in regulation allows two putts to make par on the hole.
aveputt: Average putts per round. Number of putts taken in an 18-hole round, fewer putts per round corresponds to a higher score. 
avesand: Average sand shots per round. Shots from the sand are rare (usually only 1 to 3 per round), but can have a meaningful impact on score 
pctsndsv: Sand save percentage. Calculated by dividing the total number of sand shots that hit the green and were followed by only one putt by the total number of sand shots. 

1. Load the LGPA data into the workspace.
```{r}
LGPA_data <- read.csv('Golf_Dataset.csv')
head(LGPA_data)
```
2. Perform exploratory data analysis of numerical variables compared to przrnd.
```{r, out.width = '750px', dpi=200}
par(mfrow=c(2,4), mar=c(1,1,5,1)+1)
numer_variables <- c('rounds', 'avedist','pctfrwy', 'pctgrn', 'aveputt', 'avesand', 'pctsndsv')

for (variable in numer_variables) {
  print(variable)
  plot(LGPA_data[, variable], LGPA_data$przrnd, pch=19, ylab="przrnd", main=variable)
}
```
Notes: 
- Rounds: Appears to have a curved relationship with przrnd, data should be transformed.
- Avedist: Curved relationship with przrnd, high variability
- Pctfrwy: Curved/nonlinear relationship with przrnd
- Pctgrn: Curved relationship with przrnd, lower variability
- Aveputt: Curved relationship with przrnd, aveputt negatively related to przrnd
- Avesand: Curved relationship with przrnd, negatively related
- Pctsndsv: Curved/uncelar relationship with przrnd, high variability

3. Try transformations to linearize the data.
```{r}
par(mfrow=c(2,4), mar=c(1,1,5,1)+1)
numer_variables <- c('rounds', 'avedist','pctfrwy', 'pctgrn', 'aveputt', 'avesand', 'pctsndsv')

for (variable in numer_variables) {
  print(variable)
  plot(LGPA_data[, variable], log(LGPA_data$przrnd), pch=19, ylab="przrnd", main=variable)
}
```
Notes: 
- A log transform of the response variable (przrnd) seems to work for shifting the plots to linear relationships, should be fine for all predictors.  
- pctfrwy, aveputt may need additional transformations to get linear relationships?

4. Identify other issues that might make predictors unsuitable (unbalanced data, nonlinear behavior, etc)
- Most of the variables look well balanced, most seem to have a good spread of values across the ranges. The "avesand" variable does seem to lack data towards the upper ranges of the data compared to the lower end, could think about excluding this from the analysis. 
- Do we want to change the integer objects to factors? For rounds, looks like there is a pretty continuous spread of data. Unless we want to break into ranges, probably best to leave rounds as a continuous variable rather than a factor. 
5. Process data based on exploratory analysis findings.
```{r}
# Log transform the response
logprzrnd <- log(LGPA_data$przrnd)
LGPA_data$logprzrnd <- logprzrnd

# Center the variables
predictors <- c("rounds", "avedist", "pctfrwy", "pctgrn", "aveputt", "avesand", "pctsndsv")
LGPA_data[,predictors] <- scale(LGPA_data[,predictors])
```

6a. Run LASSO to select predictors to include in the regression
https://glmnet.stanford.edu/articles/glmnet.html
```{r}
#install.packages("glmnet")
library(glmnet)

# Fit to all of the predictors in the model (exclude names)
predictors <- c("rounds", "avedist", "pctfrwy", "pctgrn", "aveputt", "avesand", "pctsndsv")
response <- c("przrnd")
fit_lasso <- glmnet(LGPA_data[,predictors], LGPA_data[,response])
plot(fit_lasso, alpha = 0, label=TRUE) # Default

# Table for predictors by number
data.frame(1:length(predictors), predictors)
#coef(fit_lasso, s = 0.1)
#for (i in length(predictors)){
#  text()
#}
```
6b. Use regularized horseshoe prior to identify coefficients to eliminate from the model.
```{r}
library("rstanarm")
library("rstantools")
library("bayesplot")
predictors <- c("rounds", "avedist", "pctfrwy", "pctgrn", "aveputt", "avesand", "pctsndsv")
response <- c("przrnd")
LGPA_data_cleaned <- LGPA_data[,c(predictors, response)]

p <- length(predictors)
n <- nrow(LGPA_data)
p0 <- 5
slab_scale <- sqrt(0.3/p0)*sd(LGPA_data_cleaned$przrnd)
global_scale <- (p0/(p-p0))/sqrt(n)
fit_horseshoe <- stan_glm(przrnd ~ ., data = LGPA_data_cleaned, refresh=0, prior = hs(global_scale=global_scale, slab_scale=slab_scale))

# Plot
library(ggplot2)
p0 <- mcmc_areas(as.matrix(fit_horseshoe), pars=vars(-'(Intercept)', -sigma), prob_outer=0.95, area_method="scaled height")
p0 <- p0 + scale_y_discrete(limits = rev(levels(p0$data$parameter)))
p0
```
Notes: Based on the horseshoe priors, rounds, pctgrn, and aveputt have the largest coefficient values. Avedist and pctfrwy both have small coefficient values compared to the others. Avesand and pctsndsv both have coefficients somewhat close to zero, so these could potentially be excluded from the analysis. 

7. Fit basic regression model using most significant coeffficients.
```{r}
fit_1 <- stan_glm(logprzrnd ~ rounds + pctgrn + aveputt, data=LGPA_data, refresh=0)
print(fit_1)
#summary(fit_1)

# Print the R2 value for the fitted model
fit_1_R2 <- round(median(bayes_R2(fit_1)), 2)
print(paste("R2:", fit_1_R2))
```

